---
output: 
      github_document:
                  pandoc_args: --webtex
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

remotes::install_github("Asymmetrical-irt/AsymmIRT")
```


# AsymmIRT: Bayesian analysis of asymmetric IRT models (RLPE, LPE) models and symmetric IRT models

<!-- badges: start -->
[![R-CMD-check](https://github.com/Asymmetrical-irt/AsymmIRT/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Asymmetrical-irt/AsymmIRT/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

AsymmIRT is an R package that provides a comprehensive toolkit for fitting, diagnosing, and interpreting asymmetric IRT models. The package currently implement bayesian analysis of asymmetric IRT models (e.g 1PLPE, 1PRLPE, LPE, RLPE) and widely known symmetric models such as 1PL and 2PL. Other asymmetric IRT models are planned to be implemented in the future.

## Installation

AsymmIRT uses [`cmdstanr`](https://mc-stan.org/cmdstanr/) (not available on CRAN). To install it, follow these instructions [Getting started with CmdStanR](https://mc-stan.org/cmdstanr/articles/cmdstanr.html) 

The development version of AsymmIRT can be installed typing:
``` r
remotes::install_github("Asymmetrical-irt/AsymmIRT")
```

## Data simulation

This is a basic example on how to simulate data from the available asymmetric and symmetric models. The package provides \texttt{print()} and \texttt{plot()} methods for examining generated datasets. The \texttt{print()} method displays summary statistics including the mean, standard deviation, and range of true parameter values. The \texttt{plot()} method produces a histogram of total scores, allowing quick assessment of score distribution characteristics.

```{r parallel_setup}

```

```{r example}
library(AsymmIRT)

n = 25      ## number of individuals
k = 20 ## number of items
model_type   = "LPE"   ## type of model to simulate data from (e.g LPE, RLPE,1PRLPE)
seed = 123  

data <- simData(n=n , k=k, seed=seed, model_type=model_type)


plot(data)
print(data)
```

## Fitting the data to a model

Fitting models using the AsymmIRT package relies on pre-compiled models, using the instantiate package, which manages the compilation and loading of the Stan models. It is important to notice that the function fit_model() accepts the same arguments as the sample method in CmdStanR.

Additionally computing metrics such as WAIC, DIC and loo can be obtained for the available models using a ’_log’ suffix. For example, specifying ‘1PLPE_log’ instead of ‘1PLPE’, the output will allow to use the methods in the package to compute these metrics.

```{r fit_setup}
iter_warmup = 200 ## number of items
iter_sampling = 200      ## number of individuals
model_type   = "1PLPE_log"   ## type of model 
seed = 123  
data = data$df 
chains = 2 ## number of chains
parallel_chains = 2 ## Running chains in parallel
```

Notice that the function fit_model() returns a list of three objects. The data list, type of model and the **output** which is a R6 CmdStanMCMC object.


```{r fit}
data <- simData(n=25 , k=20, seed=123, model_type="LPE")
fit <- AsymmIRT::fit_model(data = data$df, iter_warmup = iter_warmup, iter_sampling = iter_sampling, mod = "LPE2_log", seed = seed, chains = chains, parallel_chains = parallel_chains)
```


```{r fit_methods}
code(fit)
```

Here we can see the code of the model. In this example priors are

$$\theta_{i} \sim N(0,1)$$
$$b_{j} \sim N(0,\sqrt{2})$$
$$\lambda_{j} \sim LogNormal(0,\sqrt{0.5})$$

The model 1PLPE is:

$$p_{ij}^{1PLPE} = P(Y_{ij} = 1) = \left( \frac{\exp(\theta - b)}{1 + \exp(\theta - b) } \right)^{\lambda_{j}}$$

Now it is possible to delve into the bayesian analysis. Notice that WAIC and loo accept other arguments, for further information visit the webpage of the **loo** package.
 
```{r fit_summary}
summary(fit,variable = c("b","theta","lambda"), ci = 0.95)
```


```{r fit_dic}
DIC(fit)
```


```{r fit_waic}
WAIC(fit)
```


```{r fit_loo}
loo(fit)
```



Lets plot the item icc and item info


```{r item_icc}
item_icc(fit, item = 2, theta_lim = c(-3,3))
```


```{r item_info}
item_info(fit, item = 2, theta_lim = c(-3,3))
```

To visually diagnose the convergence of the MCMC chains, the package AsymmIRT also allows to
obtain traceplots of the parameter

```{r item_plot_trace}
plot_trace(fit, variable = "b")
```

Adittionally, to plot the HPD intervals alongside the estimation median
the user can use the plot method of the Asymmfit class

```{r hpd}
plot(fit, variable = "b")
```
